---
title: "Lab 6"
author: "Allison Gracer"
date: "11/1/2021"
output: html_document
---  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(broom)
library(equatiomatic)
```

## Example of a rank-based test 

Refer to lecture 11 for additional notes on rank-based test.
What we are comparing across groups is the ranks instead of the actual values. 

We'll make our own samples, using a pseudorandom generator. type `?set.seed` in console to read about this function. It is a random number generator. Allows us to set a starting point from which we want the random values. 

```{r}
set.seed(1414) # set.seed tell R to create a random sample of integer values
gp_1 <- sample.int(20, size = 15, replace = TRUE) # max is 20, sample size is 15, replace is true (re-sampling)
set.seed(1424)
gp_2 <- sample.int(30, size = 15, replace = TRUE) # max is 30, sample size is 15 replace is true
```

```{r}
hist(gp_1) # making a quick histogram with a vector of numbers

hist(gp_2) # sample size is small, not convinced based on central limit therorem that these means will be normally distributed

```

Try a t-test:
```{r}
t.test(gp_1, gp_2)
```

P value = .1982 
What does this p-value mean?

If these samples were drawn from populations with the same mean, the probability of taking two random samples with means *at least as different* as the sample means we found by random chance (taking into account spread, n) is 19.8% (write 19.8% using in-line code for reports).

Do we have sufficient evidence to reject the null hypothesis? No, it's super likely, 20% chance that these were

Retain(fail to reject) the null hypothesis. 

Warning: "accept" null, dont use this language in a final report. You should be saying there is or isn't significant difference between the samples. 

Now let's compare this outcome to a rank-based test.

## Mann Whitney U unpaired rank-based test (to compare ranks)

```{r}
mwu <- wilcox.test(gp_1, gp_2)

mwu # look at the outcome of mwu we ran above
```

null hypotheiss: drawn from populations with the same ranks

p-value is 0.28

what does this p-value say?

If these samples were drawn from populations with the same ranks(medians), the probability of finding two samples with ranks *at least as different* as those in our samples is 0.28.

There is no significant difference in ranks(medians) between group 1 and group 2 (statistical summary). 

Median scores for group 1 (M = median value i.e. 14) and group 2 (M = 12) did not differ significantly (Mann Whitney U test: U(df) = 86, p = 0.28).

Does this differ from what we learned from running our t-test?
Our conclusion is the same that we retain the null hypothesis of either equal means or equal medians regardless of etiher test we use. 

Lose points if: you have gone through the work to chose the correct responsbile test to use, make sure that it aligns with everything else you are presenting to your audience. 

ANOVA done to present means, and box plot representing a line only at the median. Once you decide the appropriate method, be sure you are using that same method to present it visually as well. 



