---
title: "Lab 6"
author: "Allison Gracer"
date: "11/1/2021"
output: html_document
---  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(broom)
library(equatiomatic)
```

## Example of a rank-based test 

Refer to lecture 11 for additional notes on rank-based test.
What we are comparing across groups is the ranks instead of the actual values. 

We'll make our own samples, using a pseudorandom generator. type `?set.seed` in console to read about this function. It is a random number generator. Allows us to set a starting point from which we want the random values. 

```{r}
set.seed(1414) # set.seed tell R to create a random sample of integer values
gp_1 <- sample.int(20, size = 15, replace = TRUE) # max is 20, sample size is 15, replace is true (re-sampling)
set.seed(1424)
gp_2 <- sample.int(30, size = 15, replace = TRUE) # max is 30, sample size is 15 replace is true
```

```{r}
hist(gp_1) # making a quick histogram with a vector of numbers

hist(gp_2) # sample size is small, not convinced based on central limit therorem that these means will be normally distributed

```

Try a t-test:
```{r}
t.test(gp_1, gp_2)
```

P value = .1982 
What does this p-value mean?

If these samples were drawn from populations with the same mean, the probability of taking two random samples with means *at least as different* as the sample means we found by random chance (taking into account spread, n) is 19.8% (write 19.8% using in-line code for reports).

Do we have sufficient evidence to reject the null hypothesis? No, it's super likely, 20% chance that these were

Retain(fail to reject) the null hypothesis. 

Warning: "accept" null, dont use this language in a final report. You should be saying there is or isn't significant difference between the samples. 

Now let's compare this outcome to a rank-based test.

## Mann Whitney U unpaired rank-based test (to compare ranks)

```{r}
mwu <- wilcox.test(gp_1, gp_2)

mwu # look at the outcome of mwu we ran above
```

null hypotheiss: drawn from populations with the same ranks

p-value is 0.28

what does this p-value say?

If these samples were drawn from populations with the same ranks(medians), the probability of finding two samples with ranks *at least as different* as those in our samples is 0.28.

There is no significant difference in ranks(medians) between group 1 and group 2 (statistical summary). 

Median scores for group 1 (M = median value i.e. 14) and group 2 (M = 12) did not differ significantly (Mann Whitney U test: U(df) = 86, p = 0.28).

Does this differ from what we learned from running our t-test?
Our conclusion is the same that we retain the null hypothesis of either equal means or equal medians regardless of etiher test we use. 

Lose points if: you have gone through the work to chose the correct responsbile test to use, make sure that it aligns with everything else you are presenting to your audience. 

ANOVA done to present means, and box plot representing a line only at the median. Once you decide the appropriate method, be sure you are using that same method to present it visually as well. 

do ?wilcox.test in consol to see documentation
scroll down to values for names of the value you want to write in in-line text 
type mwu$p.value in consol to pull value up and write as in-line text

Kruskal Wallace Test `kruskal.test`

## Linear regression

Simple linear regression (single dependent variable, a single independent variable)


```{r}
# make an exploratory plot of penguin body mass (y-axis) versus flipper length (x-axis)
# deal with exploratory analysis by putting include = FALSE in the code chunk header so it does not show up in your knitted document
ggplot(data = penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point() + 
  geom_smooth(method = "lm") # add a linear model, this is ok for exploratory, but do not add this to final if I cant explain it entirely

```

Find a linear regression model using ordinary least squares describing the relationship between flipper length and body mass for these penguins.

3 pieces:

- what type of model?
- What is the relationship to model (Dependent Variable DV ~ Ind. Variable IV(s))
- Where is the data that's used to create this model?

```{r}
penguin_lm <- lm(body_mass_g ~ flipper_length_mm, data = penguins) # how does body_mass_g change in relation to flipper_length, get my data from penguins, lm = linear model

#then call it in console penguin_lm to see what you got
# y= mx+b y = body mass mx = flipper length + some error
# body_mass = 49.7 (flipper_length) - 5780.83
# coefficient for flipper length 49.7 means for everyone mm increase in flipper length i expect a 49.7 gram increase in body mass because that's the slope that explains the relationship between these 2 variables. 

#penguin_lm

# Call:
# lm(formula = body_mass_g ~ flipper_length_mm, data = penguins)

# Coefficients:
      #(Intercept)  flipper_length_mm  
         # -5780.83              49.69  
```

### broom package returns model outputs as tidy data frames

```{r}
penguin_lm_tidy <- broom::tidy(penguin_lm) #from the broom package #you can do tidy(penguin_lm)

penguin_lm_tidy

broom::glance()
```

How can I actually include my model equation in a report?

In comes equatiomatic!!

```{r}
extract_eq(model = penguin_lm, use_coefs = TRUE)
```

```{r}
plot(penguin_lm)  # these diagnostic plots give us some ways to start exploring the assumptions we have when we decide to run a linear model
```



